GVN-hoist: hoisting expressions from control flow

### 2016-08-04 hoist.tex
reorganize the sections

\section{Hoisting to improve performance}
Hoisting also reduces critical path length of execution in out of
order machines (but not in sequential machines), by exposing ILP
before the conditional where the instruction was hoisted.

@@

\section{Hoisting scalars}
Scalars are the easiest to hoist because we do not have to analyze
them for aliasing memory references. As long as all the operands are
available (which in SSA form we do, or may be there are chained
dependencies), the scalar can be hoisted. Care must be taken in case
of hoisting scalars too far, as that may increase register pressure
and result in spills. For example hoisting a scalar past a call. In
that case the call may result in save and restore of the register the
scalar may be defined to. In our current implementation we do not
hoist scalars past a call. Another way to mitigate this problem is be
to reinstantiate (rematerialize) the computation after a call (may be
as a different pass).

\section{Using MemorySSA}
\subsection{Iterative update to memory ssa during the hoisting process}

\section{Without MemorySSA using the dominance information}

\section{Hoisting stores}
We need to value number the address and the value being stored.

\section{Hoisting calls}
Calls can be hoisted by value-numbering all the arguments. For safety
we have to check that there are no other side-effects between a call and
its hoisting point, just like loads and stores.

\section{Comparative analysis of different approaches to hoisting}
\subsection{Triangle approach}
This is the most basic approach to code-hoisting when the instructions are
hoisted to immediate parent in a diamond like structure. When there are
identical computations in both the branches of a conditional, and all the
legality checks are met, the instruction can be hoisted to the predecessor basic
block. This approach is simple to implement and has minimal impact on the
register pressure.

@@

\section{Hoisting bottom up with optimistic approach}
Optimistic means collect the set of all potential redundancies and
then discard the bad ones.

@@

\section{Partition the list of hoisting candidates to maximize hoisting}
\cite{sec:partition}
The set of candidates are sorted in the order of their DFS-in
numbers. So that candidates close to each other in the list are also
closer (??) in the CFG.

@@

\section{Cost models}
\subsection{In the presence of calls}
Hoisting scalars across calls is tricky because it can increase the
number of spills.

\section{Enabling LICM by code-hoisting}
loop:
if (a)
  i1;
else
  i1;

\section{Improving the Inliner heuristics}
Reducing the number of instructions also reduces the inlining cost and
hence improves the inliner heuristics.

\section{Improving vectorization}
It will also benefit vectorizer by reducing the number
of basic blocks, should all the instructions in conditional get hoisted.

>>>>

\section{Code hoisting}
Hoisting also reduces critical path length of execution in out of
order machines (but not in sequential machines), by exposing ILP
before the conditional where the instruction was hoisted.

@@

\subsection{Hoisting scalars}
Scalars are the easiest to hoist because we do not have to analyze
them for aliasing memory references. As long as all the operands are
available (which in SSA form we do, or may be there are chained
dependencies), the scalar can be hoisted. Care must be taken in case
of hoisting scalars too far, as that may increase register pressure
and result in spills. For example hoisting a scalar past a call. In
that case the call may result in save and restore of the register the
scalar may be defined to. In our current implementation we do not
hoist scalars past a call. Another way to mitigate this problem is be
to reinstantiate (rematerialize) the computation after a call (may be
as a different pass).

\subsection{Using MemorySSA}
\subsection{Iterative update to memory ssa during the hoisting process}

\subsection{Without MemorySSA using the dominance information}

\subsection{Hoisting stores}
We need to value number the address and the value being stored.

\subsection{Hoisting calls}
Calls can be hoisted by value-numbering all the arguments. For safety
we have to check that there are no other side-effects between a call and
its hoisting point, just like loads and stores.

\subsection{Cost models}

\subsection{Reduce register pressure}
Hoisting scalars across calls is tricky because it can increase the
number of spills.

\subsection{In the presence of calls}
Caller saved registers.

\subsection{Enabling LICM by code-hoisting}
loop:
if (a)
  i1;
else
  i1;

\subsection{Improving the Inliner heuristics}
Reducing the number of instructions also reduces the inlining cost and
hence improves the inliner heuristics.

\subsection{Improving vectorization}
It will also benefit vectorizer by reducing the number
of basic blocks, should all the instructions in conditional get hoisted.

\section{Experimental Evaluation}

\subsection{Comparative analysis of different approaches to hoisting}
\subsection{Triangle approach}
This is the most basic approach to code-hoisting when the instructions are
hoisted to immediate parent in a diamond like structure. When there are
identical computations in both the branches of a conditional, and all the
legality checks are met, the instruction can be hoisted to the predecessor basic
block. This approach is simple to implement and has minimal impact on the
register pressure.

@@

\subsection{Hoisting bottom up with optimistic approach}
Optimistic means collect the set of all potential redundancies and
then discard the bad ones.

@@

\subsection{Partition the list of hoisting candidates to maximize hoisting}
\cite{sec:partition}
The set of candidates are sorted in the order of their DFS-in
numbers. So that candidates close to each other in the list are also
closer (??) in the CFG.

@@

\section{Conclusion and Future Work}
